Estimating dark matter halo concentration using the integrated mass profile

### 2015-11-06 paper/paper.tex
abstract revisado

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Estimating dark matter halo concentration from the integrated mass profile}
\author[C. Poveda, J.E. Forero-Romero, J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{xxxx@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

@@

Here we show that the concentration parameter in dark matter halos from
N-body simulations can be reliably estimated from the radially integrated mass
profile.
The main advantage of this method is that it uses the full
particle information without any binning, {\bf which makes it very
  switable in cases when resolution becomes important}. We test our
method both on mock and N-body halos to compare it against two popular
methods to find concentrations: maximum radial velocity measurements
and radial particle binning to {\bf fit the density profile}. Tests on
the mock halos show that the accuracy of our method to recover input
concentrations varies with the halo {\bf resolution}. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy. {\bf For the maximum radial
  velocity method it is on the order of $20\%$ while density profile
  methods perform worst at the low resolution limit since they require
  at least 200 particles inside the virial radius to make a reliable
  estimate of the density profile}. For halos with $2\times 10^4$
particles our method achieves an accuracy of $0.01\%$ while the
velocity and density profile methods achieve $0.1\%$ and $1\%$,
respectively. We measure the mass-concentration relationship on N-body
data taking care of using halos sampled with at least $200$
particles. We find that at low masses $10^{12}\hMsun <
M<10^{13}\hMsun$ our method yields median concentration values lower
by a $15\%-20\%$ compared to the velocity and density methods ({\bf
  why?}).  At higher masses $3\times 10^{13}\hMsun < M < 2\times
10^{14}\hMsun$ the three methods give similar results.
\end{abstract}

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Estimating dark matter halo concentration using the integrated
  mass profile} 
\author[C. Poveda, J.E. Forero-Romero, J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{xxxx@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

@@

In a dark matter dominated Universe galaxy properties are driven by
dark matter halo characteristics. 
One of the key structural halo properties is the concentration. 
[EN QUE MODELOS IMPORTA LA CONCENTRACION?]
For these reasons there has been a lot of effort to have a reliable
concentration estimate from N-body simulations.
Here we show that the concentration parameter in dark matter halos from
N-body simulations can be reliably estimated from the radially integrated mass
profile.
The main advantage of this approach is that it uses the full
particle information without any binning, which makes it reliable in
cases when low numerical resolution becomes a limitation for other
methods.   
We show the good performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy. 
For the maximum radial velocity method it is on the order of $20\%$
while density profile methods perform worst at the low resolution
limit since they require at least 200 particles inside the virial
radius to make a reliable estimate of the density profile. 
For halos with $2\times 10^4$ particles our method achieves an
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We measure the mass-concentration relationship on N-body
data taking care of using halos sampled with at least $200$
particles. We find that at low masses $10^{12}\hMsun <
M<10^{13}\hMsun$ our method yields median concentration values lower
by a $15\%-20\%$ compared to the velocity and density methods ({\bf
  why?}).  
At higher masses $3\times 10^{13}\hMsun < M < 2\times
10^{14}\hMsun$ the three methods give similar results.
\end{abstract}
